---
title: "Reproducible Research, Assignment 1"
author: "Amarins van de Voorde"
date: "21-7-2020"
output: 
  html_document: 
    keep_md: yes
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The first step of the assignment is to load and process the data. In my code I have assumed that the activity.zip file has been placed in the working directory. I also converted the column "date" to type Date.
```{r Loading, reading and processing the data}
unzip("activity.zip")
activity<-read.csv("activity.csv")
summary(activity)
activity$date<-as.Date(activity$date)
```


### Mean total number of steps per day
This is a histogram of the total number of steps taken each day.
```{r Histogram total number of steps per day}
totalsteps<-aggregate(list(total.steps=activity$steps),list(date=activity$date),FUN=sum,na.rm=FALSE)
hist(totalsteps$total.steps,main="Histogram of total number of steps per day",xlab="Total number of steps per day",col="blue",breaks=10)
```

The mean and median total number of steps taken each day are, respectively:
```{r Mean and median total number of steps per day}
mean(totalsteps$total.steps,na.rm=TRUE)
median(totalsteps$total.steps,na.rm=TRUE)
```


### Average daily activity pattern
This is a time series plot of the averate number of steps taken.
```{r Average number of steps per interval}
averagesteps<-aggregate(list(mean.steps=activity$steps),list(interval=activity$interval),FUN=mean, na.rm=TRUE)
plot(mean.steps~interval,type="l",averagesteps,main="Average number of steps per interval, averaged across all days")
```

The 5-minute interval that, on average, contains the maximum number of steps, is: 
```{r Interval with maximum number of steps}
max.average.steps<-which.max(averagesteps$mean.steps)
averagesteps[max.average.steps,]
```


### Inputing missing values
In this step, the _NA_ values are replaced by the average number of steps for that interval. First we determine how many records there are that contain an _NA_ value.
```{r Determining number of NA values}
sum(is.na(activity$steps))

```

```{r Imputing the NA values}
library(dplyr)
activity.imputed<-full_join(activity,averagesteps)
activity.imputed$steps.imputed<-ifelse(is.na(activity.imputed$steps),activity.imputed$mean.steps,activity.imputed$steps)
```

This is a histogram of the total number of steps taken each day, after the missing values have been imputed. The data have first been aggregated.
```{r Imputed total number of steps per day}
totalsteps.imputed<-aggregate(list(imputed.total.steps=activity.imputed$steps.imputed),list(date=activity.imputed$date),FUN=sum)
hist(totalsteps.imputed$imputed.total.steps,main="Histogram of imputed total number of steps per day",xlab="Imputed total number of steps per day",col="blue",breaks=10)
```

The mean and mean number total number of steps taken each day are, respectively:
```{r Mean and median imputed total number of steps per day}
mean(totalsteps.imputed$imputed.total.steps)
median(totalsteps.imputed$imputed.total.steps)

```
When comparing to the mean and median total number of steps as calculated before imputing the data, we observe that the mean total number of steps per day has stayed the same. The median total number of steps per day has changed after imputing the data, and is now equal to the mean total number of steps per day.


### Are there differences in activity patterns between weekdays and weekends?
The imputed data has been categorized as weekday-data or weekend-data. The resulting dataset has been grouped by type of day, so weekend or weekday, and interval, and has been summarized by calculating the mean for each group. The resulting groups have been plotted in a panel plot.
```{r Imputed mean number of steps per interval, for weekday and weekend}
library(chron)
library(lattice)
activity.imputed$weekday<-ifelse(is.weekend(activity.imputed$date),"weekend","weekday")
activity.imputed.grouped=activity.imputed %>% group_by(weekday,interval) %>% summarise_at(vars("steps.imputed"),mean)
xyplot(activity.imputed.grouped$steps.imputed~activity.imputed.grouped$interval|activity.imputed.grouped$weekday,layout=c(1,2),main="Imputed average number of steps per interval, for weekend and weekday",xlab="Interval",ylab="Imputed average steps",type="l")
```
